### 2015-06-10 content/abstract.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

Continuous Datalog semantics

CoDeR Logical Model (minimal algebra and stream data model) for Continuous Deductive Reasoning

Physical Model and Windowed Rete Algorithm

R4 (\& Squall)

Evaluation (against Sparkwave \cite{sparkwave} and CQELS)