### 2015-06-10 content/conclusions.tex

\begin{nestedsection}{Conclusions}{conclusions}
	We are more semantically sound than other RDF Stream Processing offerings and more ``continuous'' than other continuous reasoning offerings, without sacrificing processing efficiency.

%	\subimport{conclusions/}{future-work.tex}
\end{nestedsection}