### 2015-06-10 content/introduction.tex

\begin{nestedsection}{Introduction}{intro}
	Lack of consensus on stream processing semantics within RDF stream processing community, even more dramatic between RSP and continuous reasoning communities.

	Motivating use case (Glacier stuff, to match evaluation?)
\end{nestedsection}