### 2015-06-10 template/part_locators.tex

\graphicspath{{../content/images/}}   % Location of your graphics files
\newcommand{\inputgraphics}[3][width=0.9\textwidth]{\IfFileExists{../content/images/#2}{\includegraphics[#1]{../content/images/#2}}{}} % Define a new command to input text from other files stored in the content directory.
\newcommand{\inputdoc}[2][content]{\IfFileExists{../#1/#2}{\input{../#1/#2}}{}} % Define a new command to input text from other files stored in the content directory.
\newcommand{\includechap}[2][content]{\IfFileExists{../#1/#2}{\include{../#1/#2}}{}} % Define a new command to include tex from other files stored in the content directory.