### 2015-06-10 content/personal-details-ieee.tex
Added the IOS article style (I think this is the right one, though the call for papers section doesn't mention the call I thought we were working towards...), and migrated preambles and personal-details files to be for specific article styles

% Personal Details
\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}
\begin{IEEEkeywords}
      Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation
\end{IEEEkeywords}
%\keywords   {Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation}
\author    {\IEEEauthorblockN{David L. Monks, Rehab ???, Dr. Nicholas M. Gibbins}
            \IEEEauthorblockA{
                              Web and Internet Science\\
                              Electronics and Computer Science\\
                              Physical Sciences and Engineering\\
                              University of Southampton\\
                              \{dm11g08,r???,nmg\}@ecs.soton.ac.uk
                              }
            }
% TODO fix Rehab's details
\date{\today}