### 2015-06-12 content/personal-details-els.tex
Added the CORRECT article class, and set up the template correctly apart from the problem listed below.  Also fixed the details for Rehab in all the personal-details-*.tex
PROBLEMS
- inductive step still overflows both column and page
- abstract or the rest of the frontmatter does not format properly in twocolumn mode

% Personal Details
\begin{frontmatter}
	\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}

	\author{David L. Monks}
	\ead{dm11g08@ecs.soton.ac.uk}
	\author{Rehab Albeladi}
	\ead{raab1g09@ecs.soton.ac.uk}
	\author{Nicholas M. Gibbins}
	\ead{nmg@ecs.soton.ac.uk}
	% \ead{\{dm11g08,raab1g09,nmg\}@ecs.soton.ac.uk}
	\address{
	          Department of Electronics and Computer Science,\\
	          Faculty of Physical Sciences and Engineering,\\
	          University of Southampton,\\
	          University Rd., Southampton, SO17 1BJ,\\
	          United Kingdom
	}
	% TODO fix Rehab's details

	\begin{abstract}
		\input{abstract.tex}
	\end{abstract}

	\begin{keyword}
	      Deductive Reasoning\sep Streamed Data\sep Semantic Data\sep Continuous Processing\sep Plan Optimisation
	\end{keyword}

	\date{\today}

\end{frontmatter}