### 2015-06-29 content/personal-details-ios.tex

% Personal Details
\begin{frontmatter}
	\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}
	\runningtitle{CoDeR Model and R4 Algorithm: Continuous Deductive Reasoning over Streamed RDF Data}

	\review{}{}{}

	\author{\fnms{David L.} \snm{Monks}}
	\sep
	\author{\fnms{Rehab} \snm{Albeladi}}
	and
	\author{\fnms{Nicholas} \snm{Gibbins}}
	\runningauthor{D.L. Monks et al.}
	\address{
	          Department of Electronics and Computer Science,\\
	          Faculty of Physical Sciences and Engineering,\\
	          University of Southampton,\\
	          University Rd., Southampton, SO17 1BJ,\\
	          United Kingdom\\
	          E-mail: \{dm11g08,raab1g09,nmg\}@ecs.soton.ac.uk
	}
	% TODO fix Rehab's details

	\begin{abstract}
		\input{abstract.tex}
	\end{abstract}

	\begin{keyword}
	      Deductive Reasoning\sep Streamed Data\sep Semantic Data\sep Continuous Processing\sep Plan Optimisation
	\end{keyword}

	\date{\today}

\end{frontmatter}