### 2015-07-14 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
fixed application notation and bracketing error in inductive proof

\begin{nestedappendix}{Entailment Semantics: Proof of Inductive Lemma}{semantics: proof for entailment}
	\begin{figure*}
		\centering
		\caption{General Case: expresses the set difference $W^{P,x}_{i} \setminus W^{N,y}_{i}$ in terms of windows of range 1, eliminating $y$ as a variable.}
		\labelfig{semantics: induction general case}
		\begin{align*}
			\mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) & = W^{P,x}_{i} \setminus W^{N,y}_{i} \\
			& = \mathop{\cup}^{x-1}_{d=0} W^{P,1}_{i-d} \setminus \mathop{\cup}^{y-1}_{j=0} W^{N,1}_{i-j} && \text{by Axiom~\ref{axiom:continuous datalog: window composition}} \\
			& = \mathop{\cup}^{x-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{y-1}_{j=0} W^{N,1}_{i-j} \right) \\
			& = \mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{y-1}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
			& = \mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{d}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) && \text{by Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } y \geq x
		\end{align*}
	\end{figure*}

	\begin{figure*}
		\centering
		\caption{Base Case: proves Lemma~\ref{hyp:continuous datalog: entailment lemma} for the case ${x = 1}$.}
		\labelfig{semantics: induction base case}
		\begin{align*}
			E^T_{i} \setminus \mathop{\cap}^{1}_{d=1} E^T_{i-d} & = W^{P,1}_{i} \setminus W^{N,y}_{i} \\
			& = W^{P,1}_{i} \setminus W^{N,1}_{i} && \text{by \textbf{General Case}} \\
			& = \left( E^T_{i} \setminus E^T_{i-1} \right) \setminus \left( E^T_{i-1} \setminus E^T_{i} \right) && \text{by Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} \\
			& = E^T_{i} \setminus \left( E^T_{i-1} \setminus \left( E^T_{i-1} \setminus E^T_{i} \right) \right) \\
			& = E^T_{i} \setminus \left( E^T_{i-1} \cap E^T_{i} \right) \\
			& = \left( E^T_{i} \setminus E^T_{i-1} \right) \cup \left( E^T_{i} \setminus E^T_{i} \right) \\
			& = E^T_{i} \setminus E^T_{i-1} \\
			& = E^T_{i} \setminus \mathop{\cap}^{1}_{d=1} E^T_{i-d} && \qedhere
		\end{align*}
	\end{figure*}

	\begin{figure*}
		\centering
		\caption{Inductive Case: proves Lemma~\ref{hyp:continuous datalog: entailment lemma} for the case ${x = k + 1}$, assuming that it holds for ${x = k}$.}
		\labelfig{continuous datalog: entailment lemma inductive case}
		\begin{align*}
			E^T_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E^T_{i+1-d} & = W^{P,k+1}_{i+1} \setminus W^{P,k+1}_{i+1} \\
			& = \mathop{\cup}^{k-1}_{d=-1}\mathop{\cap}^{d}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) && \text{by \textbf{General Case}} \\
			& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{i-j} \setminus W^{N,1}_{i-j} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{d}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
			& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{d}_{j=-1} W^{N,1}_{i-j} \right) \\
			& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{d}_{j=0} W^{N,1}_{i-j} \setminus W^{N,1}_{i+1} \right) \\
			& = \left( W^{P,1}_{i+1} \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{d}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \right) \setminus W^{N,1}_{i+1} \\
			& = \left( W^{P,1}_{i+1} \cup \left( W^{P,k}_{i} \setminus W^{N,k}_{i} \right) \right) \setminus W^{N,1}_{i+1} \\
			& = \left( \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) &&
				\begin{aligned}
					&\text{by \textbf{Inductive Step}} \\
					&\text{and Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}}
				\end{aligned} \\
			& = \left( E^T_{i+1} \setminus E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \\
			& = \left( E^T_{i+1} \setminus E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \cup \left( E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
			& = \left( E^T_{i+1} \setminus \left( E^T_{i} \cap E^T_{i+1} \right) \right) \cup \left( \left( E^T_{i} \cap E^T_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
			& = \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( \left( E^T_{i} \cap E^T_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
			& = \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( E^T_{i+1} \cap \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
			& = \left( E^T_{i+1} \cap \left( E^T_{i} \right)^C \right) \cup \left( E^T_{i+1} \cap \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right)\\
			& = E^T_{i+1} \cap \left( \left( E^T_{i} \right)^C \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
			& = E^T_{i+1} \setminus \left( E^T_{i} \setminus \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
			& = E^T_{i+1} \setminus \left( E^T_{i} \cap \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
			& = E^T_{i+1} \setminus \mathop{\cap}^{k}_{d=0} E^T_{i-d} \\
			& = E^T_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E^T_{i+1-d} && \qedhere
		\end{align*}
	\end{figure*}
\end{nestedappendix}