### 2015-07-14 content/personal-details-ieee.tex

% Personal Details
\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}
\begin{IEEEkeywords}
      Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation
\end{IEEEkeywords}
%\keywords   {Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation}
\author    {\IEEEauthorblockN{David L. Monks, Rehab Albeladi, Nicholas Gibbins}
            \IEEEauthorblockA{
                              Electronics and Computer Science\\
                              University of Southampton\\       
                              Southampton, SO17 1BJ\\
                              United Kingdom \\
                              \{dm11g08,raab1g09,nmg\}@ecs.soton.ac.uk
                              }
            }
% TODO fix Rehab's details
\date{\today}

\maketitle

\begin{abstract}
	\input{abstract.tex}
\end{abstract}