### 2015-07-15 article.tex
TESTED and WORKING: re-organised folder to work without latexbuild.sh, instead running off "$ make", and possible to run easily using "$ pdflatex article.tex" and "$ bibtex article".
Also got rid of unused images and tex files

\documentclass[twocolumn,preprint,3p,number]{template/elsarticle}
% \documentclass[preprint,5p,number]{elsarticle}

\journal{Web Semantics}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amscd}

\usepackage{hyperref}

\usepackage{subfig}
\graphicspath{{images/}}
\DeclareGraphicsExtensions{.pdf} % ,.jpeg,.jpg,.png}

\theoremstyle{plain}
\newtheorem{thrm}{Theorem}
\newtheorem{lem}{Lemma}[thrm]
\newtheorem{axiom}{Axiom}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

% Document relocation and referencing
\usepackage{import}
\usepackage{template/labelledrefs}

\include{template/ebnf}
\include{template/outerjoins}

\makeatletter
\def\@author#1{\g@addto@macro\elsauthors{\normalsize%
    \def\baselinestretch{1}%
    \upshape\authorsep#1\unskip\textsuperscript{%
      \ifx\@fnmark\@empty\else\unskip\sep\@fnmark\let\sep=,\fi
      \ifx\@corref\@empty\else\unskip\sep\@corref\let\sep=,\fi
      }%
    \def\authorsep{\unskip,\space}%
    \global\let\@fnmark\@empty
    \global\let\@corref\@empty  %% Added
    \global\let\sep\@empty}%
    \@eadauthor={#1}
}
\makeatother

\begin{document}

	\subimport{content/}{personal-details-els.tex}

	\subimport{content/}{introduction.tex}
	% \subimport{../content/}{preliminaries.tex}
	\subimport{content/}{contribution.tex}
	\subimport{content/}{evaluation.tex}
	\subimport{content/}{conclusions.tex}
%	\subimport{../content/}{future-work.tex}

%	\subimport{../content/}{appendices.tex}
  \section*{References}
	\bibliographystyle{template/elsarticle-harv}
	\bibliography{bibliography}
\end{document}